### 2020-03-19 ChapterProblemDefinition/chapterProblemDefinition.tex
Fixed: Proof-read of Problem Definition

In this chapter the requirements of a solution to scalable real-time physics and the challenges of distributing a real-time physics engine will be discussed, using a naive approach as an example. Some potential solutions to distributing real-time physics and their trade-offs will be considered. Finally, this study's proposed solution, AP, (originally presented in \cite{brown2019aura}) will be described, along with high-level overviews of the challenges with this approach and the implemented solutions for the solved problems and proposed solutions for the remaining problems.

@@

%TODO: Discuss scalability
In general terms, distributing work between servers have the following overheads: the logic of determining messages to be sent; the processing overheard of sending the messages themselves; and the logic of handling received messages from other servers.
%For example, in a 
To achieve the greatest scalability, a solution must minimise overheard while maximising useful work done by each server.

The types of errors that can occur in distributed real-time physics simulations will now be discussed. 
Types of errors include:
\begin{itemize}
	\item Time-space inconsistencies, in which two different objects occupy the same space at the same time, but on different servers.
	\item Two objects occupying the same space at the same time, which can also lead to late collisions and missed collisions. 
	\item Diverging results and object state conflict, in which the same object has different states on different servers, which can be caused by collision response results differing on different servers.
	\item The same object having different states on different servers can lead to objects `jittering', as the object state is conflicting between the two servers. 
	\item Objects not being simulated (missed updates), which could occur when objects are migrating between servers. 
	\item Objects being duplicated and objects being lost. 
\end{itemize}

@@

@@

\subsection{Object state conflict}
The challenges involved with two servers reaching a consensus on the state of an object if the two objects have conflicting states will now be discussed. If a solution is used in which two servers simulate the same object (such as in an overlapping region between servers), diverging results from collisions can result in the same object can end up in different states on different servers. These conflicts in state may be small, in which case they can be resolved by either choosing the state from one server or taking the average of the two objects, however, this may appear as `jitter' if the differences between states is large enough. Furthermore, cases may exist when the two states have diverged significantly, such as the stack of cubes example in \ref{divergingResults}, in which case averaging of the two states cannot be used. One server's state would have to be chosen, which would appear on the other server as though the objects have `teleported' (instantaneously changed position). In addition to this, prior to the server having its object's states overwritten, those objects may have affected other objects in only that server. For example, if we continue with the stack of cubes example in \ref{divergingResults}, suppose the stack of cubes fell over on one server and one of the cubes collided with a sphere, that is only being simulated by that sever (for example, if it is outside of the overlapping region). The sphere is given a velocity as a result of the collision. If the cubes in the stack are then updated to be the same as on a remote server, essentially rewriting the history of the cubes, the end result would be a stack of cubes that is upright and sphere with a velocity from a collision that has been effectively erased from the simulation history.

If objects that are going to collide are migrated between servers, they need to be migrated and arrive before a collision occurs. If the arrival takes place late, this may result in the two objects overlapping significantly (occupying the same space at the same time). The consequence of these late collisions are unstable collision response. To understand how late collisions lead to unstable collision response, we must first understand how in real-time collision detection works. In real-time physics engines, objects move in discrete steps, as a result of this, when two objects collide, they overlap each other i.e. penetrate. The collision is resolved by calculating the forces resulting from the collision and moving the objects so they no longer overlap (in PhysX if penetration is high, the latter may be done over several physics time-steps). If the penetration is high, such as in the case of a late collision response or if two objects are moving at high speed towards one another, the collision response can no longer guarantee stable results.

\subsection{Missed collisions}
In cases of higher latency, or small or fast moving objects, collisions may be missed entirely. Similar to the bullet-through paper problem (where the discrete updating of objects means the two objects that should collide, never overlap and trigger a collision response), high latency can mean the state of a remote object can mean the state of the object is received so late that the local object that would collide, misses the object entirely. The larger or slower moving the objects involved are, the less likely this situation is to occur. Alternatively, if objects are migrated between servers, the object must arrive before the collisions happens, as discussed in \ref{lateCollisions}, if a delay in migration is severe enough, the collisions will be missed entirely.

\subsection{Missed updates}
Objects not being simulated (missed updates), can occur when objects are migrating between servers. When objects are being migrated between servers, they don't exist on any server, as a result the position and rotation of the object will not be updated. This would appear as though the object has frozen, how noticeable this would be to an observer would depend on the situation and the number of updates missed. For example, if the object is stationary or moving slowly, several time steps could be missed without be noticeable. Alternatively, in a scenario where a large object was in free-fall, very few update steps could be missed before an observer would notice that the object is not being simulated correctly.

\subsection{Object duplication or loss}
Any solution in which objects are migrated between servers is vulnerable to object duplication or loss. Networks do not provide absolute guarantee of message deliveries or order of message arrival. As a consequence of this, network messages for the migration of objects may be lost, in which case the object could be lost, if one server send a migrate message for an object and deletes the object from its simulation. Alternatively, a server may not delete the migrating object until an acknowledgement is received from the server it is migrating the object to, in which case, the acknowledgement may be lost and the object kept on the sending server, despite a duplicate of the object now existing on the receiving server. %Messages may arrive out of order

@@

\section{Naive Approach}\label{CrossBoundaryInteraction}
To provide a comparison and to demonstrate our contribution to the field, a naive approach is described first. The naive approach is used to highlight the challenges faced when solving this problem.

@@

Accomplishing a transfer with a simple message from one server to another is charged with many problems: (1) the message may be lost and the object disappears; (2) the message may be delayed and the object disappears and reappears; (3) there may be more than one candidate server for hosting (if object travelling quickly) resulting in duplication of an object; (4) the server an object leaves may be unaware of when it should stop simulating the object (resulting in duplicated objects)\cite{High-performanceServerSystems}; (5) an object may be colliding with another object across the boundary which would be lost if both objects are hosted on different servers (objects may pass through each other).

For additional clarity consider objects A and B in Fig. \ref{AuraProj}. Each server is aware of the objects they host but not other objects. As such, objects A and B would not interact and not collide with each other, creating a time-space inconsistency. A time-space inconsistency is defined as two objects occupying the same space at the same time. If Object B has a velocity directed towards the boundary, it would continue at its initial velocity until it traversed the boundary, at which point it would migrate to server 0. Upon completion of the migration, objects A and B would overlap far more than would be expected in a typical simulation, resulting in inappropriate collision response.

@@

\subsection{Both servers simulate objects in an overlapping region}
Simulate objects overlapping the boundary and objects interacting with those on both servers, e.g. by using an overlapping region. Object may interact with objects not in the overlapping region, so the results would diverge or the overlapping region would have to dynamically change in order to keep all interacting objects within the overlapping region. Even if all objects interacting were in the overlapping region, results would still diverge due to non-determinism of physics engines. The divergence of results due to non-determinism could be corrected by periodically syncing the objects (ensuring the objects have the same state on both servers), however, this would appear as jitter as objects would be moved from one position to their corrected position instantaneously. In addition, this would require additional network overhead for the syncing of objects between servers and extra computational overhead of needing to simulate objects twice.

@@

For new objects entering the overlapping region, there will be a delay before the new object appears on both servers, network delays would affect this and could lead to the two simulations diverging significantly. For example, if there are two fast moving objects on a collision course with each other that are expected to collide in the overlapping region. Due to network delays, each server only becomes aware of the other object once the objects have already passed each other, as a result the collision is missed entirely and the result is incorrect compared to a single simulation.
%For example, a newly entered object from one server knocked over a stack of cubes in the overlapping region, if there was a long delay in communication, the receiving server might one physics timestep be unaware of any collision, the next timestep it would update to many timesteps after the collision

\subsection{Over network object collisions}
Objects in the overlapping region interact over the network. This would be done with the two physics in lock-step, i.e. neither progresses until all computation divided between the servers has been completed for that physics step. Servers could divide the computation of collision detection and/or response between servers and then exchange results. This solution would have to simulate objects twice and would require multiple round-trips of messages over the network per physics time-step, but the collision response for objects would be identical on both servers. This approach would be affected by network delays significantly, as not only are multiple round-trips of messages required between server each physics time-step, the simulation won't advance until all messages for that physics time-step have been completed, meaning the frequency of the simulation would be negatively affected.

@@

\subsection{Remote objects as static}
Treat objects being simulated on another server as static. However, this would not lead to correct collision response, if compared to a single simulation. Momentum is conserved in collisions and no momentum is transferred to static objects, therefore the resulting momentum on the colliding object would not be the same as if the two objects were dynamic (not static) on a single simulation.
Half of all collisions would be missed, as collisions would be resolved for the object not being treated as static, but on the remote server, the collision never appears to happen, as the collision has already been resolved when receiving the updated state of the interacting object.
This could also lead to jitter, if two objects are in contact (i.e. colliding over multiple frames, such as a box stacked on top of another box), the server in which the collision is resolved on could be different every physics step. An cube stacked on top of a static cube may have a different outcome to a cube with a static cube stacked on top of it.
% It is probably more complicated than this, as physics engines do a lot to deal with objects in contact, constantly moving a static object is probably going to interfere with this.

@@

This approach would be affected by network delays. For example, using the stacked cubes example. Say the two cubes are free-falling. If the bottom cube was being simulated on a different server, the position of the bottom cube would only be updated whenever a message was received from the remote server to do so. If there was a long delay between messages from the remote server, for example, from a spike in latency or a sudden big loss in packets, the cube would not be updated and appear to have stopped falling. As the bottom cube has stopped on the receiving server, the top cube remains stationary, as it is in contact with the bottom cube. On the remote server, if messages are being received, the cube would be stationary and appear to be floating, as from its server's perspective, it is in contact with the bottom cube. As a result, both servers have incorrect states compared to a single simulation.

\subsection{All object interactions on same server}
Simulate all objects that are interacting with each other on the same server. Auras could be used to determine which objects are potentially interacting. Auras of objects would need to be exchanged. Solution would need to manage number of auras exchanged to reduce this overhead.

\section{Proposed Solution}
% Introduce this in context of above section
We term our approach Aura Projection (AP). AP tackles the problem of maintaining consistency across regions while ensuring timeliness of a simulation. In particular, AP provides the building blocks of a scalable solution to server-side physics simulations by handling all configurations of boundary cases in migratory approaches.

>>>>

In this chapter the requirements of a solution to scalable real-time physics and the challenges of distributing a real-time physics engine will be discussed. A naive approach will be used as an example. Some potential solutions to distributing real-time physics and their trade-offs will be considered. Finally, this study's proposed solution, AP, (originally presented in \cite{brown2019aura}) will be described, along with high-level overviews of the challenges with this approach and the implemented solutions for the solved problems and proposed solutions for the remaining problems.

@@

%TODO: Discuss scalability
In general terms, distributing work between servers has the following overheads: the logic of determining messages to be sent; the processing overheard of sending the messages themselves; and the logic of handling received messages from other servers.
%For example, in a 
To achieve the greatest scalability, a solution must minimise these overheads while maximising useful work done by each server.

The types of errors that can occur in distributed real-time physics simulations will now be discussed. 
Types of errors include:
\begin{itemize}
	\item Time-space inconsistencies, in which two different objects occupy the same space at the same time, but on different servers.
	\item Late collisions and missed collisions, in which after an object migration prior time-space inconsistencies lead to collisions either being detected late or missed entirely.
	\item Object state conflict, in which the same object has different states on different servers. This can lead to objects appearing to `jitter' and can be the result of diverging results of collisions between the two servers.	
%	\item Diverging results and object state conflict, in which the same object has different states on different servers, which can be caused by collision response results differing on different servers.s can lead to objects `jittering', as the object state is conflicting between the two servers. 
	\item Objects not being simulated (missed updates), which could occur when objects are migrating between servers. 
	\item Objects being duplicated and objects being lost. 
\end{itemize}

@@

If objects that are going to collide are migrated between servers, they need to be migrated and arrive before a collision occurs. If the arrival takes place late, this may result in the two objects overlapping significantly (occupying the same space at the same time). The consequence of these late collisions are unstable collision response. To understand how late collisions lead to unstable collision response, how real-time collision detection works must first be understood. In real-time physics engines, objects move in discrete steps, as a result of this, when two objects collide, they overlap each other i.e. penetrate. The collision is resolved by calculating the forces resulting from the collision and moving the objects so they no longer overlap (in PhysX if penetration is high, the latter may be done over several physics time-steps). If the penetration is high, such as in the case of a late collision response or if two objects are moving at high speed towards one another, the collision response can no longer guarantee stable results.

\subsection{Missed collisions}
In cases of higher latency, or small or fast moving objects, collisions may be missed entirely. A typical example of this is the `bullet-through paper problem', where the discrete updating of objects means that a fast moving object, the bullet, never overlaps a thin object, the paper, and a collision is never detected. The distance the bullet is updated each step is wider than the collision volume of the paper. High latency can mean the position of a remote object is received so late that it misses the overlap with a local object that would have collided entirely. The larger or slower moving the objects involved are, the less likely this situation is to occur. Alternatively, if objects are migrated between servers, the object must arrive before the collisions happens, as discussed in \ref{lateCollisions}, if a delay in migration is severe enough, the collisions will be missed entirely.

@@

\subsection{Object state conflict}
The challenges involved with two servers reaching a consensus on the state of an object if the two objects have conflicting states will now be discussed. If a solution is used in which two servers simulate the same object (such as in an overlapping region between servers), diverging results from collisions can result in the same object can end up in different states on different servers. These conflicts in state may be small, in which case they can be resolved by either choosing the state from one server or taking the average of the two objects, however, this may appear as `jitter' if the differences between states is large enough. `Jitter' refers to objects appearing to instantaneously snap back and forth between positions. Furthermore, cases may exist when the two states have diverged significantly, such as the stack of cubes example in \ref{divergingResults}, in which case averaging of the two states cannot be used. One server's state would have to be chosen, which would appear on the other server as though the objects have `teleported' (instantaneously changed position). In addition to this, prior to the server having its object's states overwritten, those objects may have affected other objects in only that server. For example, continuing with the stack of cubes example in \ref{divergingResults}, suppose the stack of cubes fell over on one server and one of the cubes collided with a sphere, that is only being simulated by that sever (for example, if it is far from the boundary). The sphere is given a velocity as a result of the collision. If the cubes in the stack are then updated to be the same as on a remote server, essentially rewriting the history of the cubes, the end result would be a stack of cubes that is upright and sphere with a velocity from a collision that has been effectively erased from the simulation history.

\subsection{Missed updates}
Objects not being simulated (missed updates), can occur when objects are migrating between servers. When objects are being migrated between servers, they don't exist on any server, as a result the position and rotation of the object will not be updated. This would appear as though the object has frozen, how noticeable this would be to an observer would depend on the situation and the number of updates missed. For example, if the object is stationary or moving slowly, several time steps could be missed without being noticeable. Alternatively, in a scenario where a large object was in a fast free-fall, very few update steps could be missed before an observer would notice that the object is not being simulated correctly.

\subsection{Object duplication or loss}
Any solution in which objects are migrated between servers is vulnerable to object duplication or loss. Networks do not provide absolute guarantee of message deliveries or order of message arrival. As a consequence of this, network messages for the migration of objects may be lost, in which case the object could be lost. For example, if one server sends a migrate message for an object and deletes the object from its simulation and the message is lost and the object never gets received and created on the receiving server. Alternatively, a server may not delete the migrating object until an acknowledgement is received from the server it is migrating the object to, in which case, the acknowledgement may be lost and the object kept on the sending server, despite a duplicate of the object now existing on the receiving server. %Messages may arrive out of order

@@

\section{Naive Approach}\label{CrossBoundaryInteraction}
To provide a comparison and to demonstrate this study's contribution to the field, a naive approach will now be described. The naive approach is used to highlight the challenges faced when solving the problem of distributed real-time physics.

@@

Accomplishing a transfer with a simple message from one server to another is charged with many problems: (1) the message may be lost and the object disappears; (2) the message may be delayed and the object disappears and reappears; (3) there may be more than one candidate server for hosting (if object travelling quickly) resulting in duplication of an object; (4) the server an object leaves may be unaware of when it should stop simulating the object (resulting in duplicated objects)\cite{High-performanceServerSystems}; (5) an object may be colliding with another object across the boundary and the collision would be missed if both objects are hosted on different servers (objects may pass through each other).

For additional clarity consider objects A and B in Fig. \ref{AuraProj}. Each server is aware of the objects they host but not other objects. As such, objects A and B would not interact and not collide with each other, creating a time-space inconsistency. A time-space inconsistency is defined as two objects occupying the same space at the same time. If Object B has a velocity directed towards the boundary, it would continue at its initial velocity until it traversed the boundary, at which point it would migrate to server 0. Upon completion of the migration, objects A and B would overlap far more than would be expected in a typical simulation, resulting in a late and inappropriate collision response.

@@

\subsection{Both servers simulate objects in an overlapping region}
Simulate objects overlapping the boundary and objects interacting with those on both servers, e.g. by using an overlapping region. Object may interact with objects not in the overlapping region, so the results would diverge or the overlapping region would have to dynamically change in order to keep all interacting objects within the overlapping region. Even if all objects interacting were in the overlapping region, results would still diverge due to non-determinism of real-time physics engines. The divergence of results due to non-determinism could be corrected by periodically synchronising the objects (ensuring the objects have the same state on both servers), however, this would appear as jitter as objects would be moved from one position to their corrected position instantaneously. In addition, this would require additional network overhead for the syncing of objects between servers and extra computational overhead of needing to simulate objects twice.

@@

For new objects entering the overlapping region, there will be a delay before the new object appears on both servers, network delays would affect this and could lead to the two simulations diverging significantly. For example, if there are two fast moving objects on a collision course with each other that are expected to collide in the overlapping region. Due to network delays, each server only becomes aware of the other object once the objects have already passed each other, as a result the collision is missed entirely and the result is incorrect compared to a centralised simulation.
%For example, a newly entered object from one server knocked over a stack of cubes in the overlapping region, if there was a long delay in communication, the receiving server might one physics timestep be unaware of any collision, the next timestep it would update to many timesteps after the collision

\subsection{Over network object collisions}
Objects in the overlapping region interact over the network. This would be done with the two physics in lock-step, i.e. neither progresses until all computation divided between the servers has been completed for that physics step. Servers could divide the computation of collision detection and/or response between servers and then exchange results. This solution would have to simulate objects twice and would require multiple round-trips of messages over the network per physics time-step, but the collision response for objects would be identical on both servers. This approach would be affected by network delays significantly, as not only are multiple round-trips of messages required between server each physics time-step, the simulation won't advance until all messages for that physics time-step have been completed, meaning the frequency of the simulation would be negatively affected. This solution also means the entire simulation can only run as fast as the slowest node as all nodes need to wait for the slowest node to complete each step of the simulation.

@@

\subsection{Remote objects as static}
Treat objects being simulated on another server as static. This would not lead to correct collision response, if compared to a centralised simulation. Momentum is conserved in collisions and no momentum is transferred to static objects, therefore the resulting momentum on the colliding object would not be the same as if the two objects were dynamic (not static) on a single simulation.
Half of all collisions would be missed, as collisions would be resolved for the object not being treated as static, but on the remote server, the collision never appears to happen, as the collision has already been resolved when receiving the updated state of the interacting object.
This could also lead to jitter, if two objects are in contact (i.e. colliding over multiple frames, such as a box stacked on top of another box), the server in which the collision is resolved on could be different every physics step. A cube stacked on top of a static cube may have a different outcome to a cube with a static cube stacked on top of it.
% It is probably more complicated than this, as physics engines do a lot to deal with objects in contact, constantly moving a static object is probably going to interfere with this.

@@

This approach would be affected by network delays. For example, using the stacked cubes example. Say the two cubes are free-falling. If the bottom cube was being simulated on a different server, the position of the bottom cube would only be updated whenever a message was received from the remote server to do so. If there was a long delay between messages from the remote server, for example, from a spike in latency or a sudden big loss in packets, the cube would not be updated and appear to have stopped falling. As the bottom cube has stopped on the receiving server, the top cube remains stationary, as it is in contact with the bottom cube. On the remote server, if messages are being received, the cube would be stationary and appear to be floating, as from its server's perspective, it is in contact with the bottom cube. As a result, both servers have incorrect states compared to a centralised simulation.

\subsection{All object interactions on same server}
Simulate all objects that are interacting with each other on the same server. Auras could be used to determine which objects are potentially interacting. Auras of objects would need to be exchanged. Solution would need to manage the number of auras exchanged to reduce this overhead.

\section{Proposed Solution}
% Introduce this in context of above section
We term this study's approach Aura Projection (AP). AP tackles the problem of maintaining consistency across regions while ensuring timeliness of a simulation. In particular, AP provides the building blocks of a scalable solution to server-side physics simulations by handling all configurations of boundary cases in migratory approaches.
